Stool Microbiome Analysis

### 2022-01-21 Rmarkdown/analysis/Fecal_Microbiome_TMAO_Analysis_211105.Rmd
Update the workflow with PSOtmao_NA. Add family level analysis.

@@

```{r Load phyloseq object}
load("../../data/processed/microbiome/phyloseq_objects/PSOtmao_211108.RData") #PSOtmao
```

@@

# Save
tiff("../../Outputs/plots/microbiome/RelAbund_Fam_TMAOTertiles_2111.tiff", width = 4, height = 4, units = 'in', res = 300)
plot.composition.relAbun
dev.off()
```

@@

```{r DESeq2 analysis - TMAO median}
#~~~~~~~~~~~~~~~
# DESeq analysis
# TMAO median
#~~~~~~~~~~~~~~~
mb <- phyloseq_to_deseq2(PSOtmao1G, ~ sex + age + bmi_final.x + tmao_mdn)
mb.ratio <- DESeq(mb, test = "LRT", reduced = ~ sex + age + bmi_final.x) # remove variable of interest here, TMAO
res.ratio <- results(mb.ratio)
alpha = 0.05
alpha.big = 0.5
sigtab = res.ratio[which(res.ratio$padj < alpha.big), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(PSOtmao1G)[rownames(sigtab), ], "matrix"))
sigtab$family_genus <- paste0(sigtab$Family, sigtab$Genus)
head(sigtab)

@@

No results when we look at TMAO median and control for age, sex, and BMI.

@@

# Save
tiff("../../Outputs/plots/microbiome/DESeq2_Genus_TMAOTertile1v3_2112.tiff", width = 8, height = 3, units = 'in', res = 300)
DESeqPlot
dev.off()
```
Save

```{r save deseq2 tmao tertile results}
write.csv(sigtabgen, "../../Outputs/tables/Microbiome/DESeq2_PSOtmao1G_TMAOTertile.csv")
```

@@

```{r save beta matrices}
save(PCoA_WU, file = "../../data/processed/microbiome/diversity/PSOtmao1G_PCoA_WU.RData")
save(PCoA_UnWU, file = "../../data/processed/microbiome/diversity/PSOtmao1G_PCoA_UnWU.RData")
save(PCoA_BC, file = "../../data/processed/microbiome/diversity/PSOtmao1G_PCoA_BC.RData")

save(dismax_wu, file = "../../data/processed/microbiome/diversity/PSOtmao1G_dismax_WU.RData")
save(dismax_uwu, file = "../../data/processed/microbiome/diversity/PSOtmao1G_dismax_UWU.RData")
save(dismax_bc, file = "../../data/processed/microbiome/diversity/PSOtmao1G_dismax_BC.RData")
```

@@

```{r Load bdiversity calculations}
load(file = "../../data/processed/microbiome/diversity/PSOtmao1G_PCoA_WU.RData")
load(file = "../../data/processed/microbiome/diversity/PSOtmao1G_PCoA_UnWU.RData")
load(file = "../../data/processed/microbiome/diversity/PSOtmao1G_PCoA_BC.RData")

load(file = "../../data/processed/microbiome/diversity/PSOtmao1G_dismax_WU.RData")
load(file = "../../data/processed/microbiome/diversity/PSOtmao1G_dismax_UWU.RData")
load(file = "../../data/processed/microbiome/diversity/PSOtmao1G_dismax_BC.RData")

@@

# ADONIS (perMANOVA) analysis
adonis2(dismax_wu ~ tmao_log + sex + age + bmi_final.y, data=df, method = "wunifrac", perm=999) # .025
#adonis2(tmao_log ~ dismax_wu + sex*age + cystatinc_bd1, data=df, method = "wunifrac", perm=999)

@@

#~~~~~~~~~~~
# BrayCurtis
#~~~~~~~~~~~
# Beta dispersion
bdis <- betadisper(dismax_bc, df$tmao_mdn)
permutest(bdis) # NS, good
boxplot(bdis, las = 2)

@@

# Unweighted
betaplot_UWU <- plot_ordination(PSOtmao1G, PCoA_UnWU, 
                    color="tmao_mdn",
                    shape = NULL,
                    axes = 1:2) +
  #scale_color_brewer(palette = "Dark2", labels = c("Below", "Above")) +
  stat_ellipse(linetype = 1) +
  labs(color = "TMAO Median Threshold") +
  theme_minimal() + 
  theme(axis.text = element_text(size = 12), # x and y axis text size
        axis.title = element_text(size = 12), # x and y label text size
        legend.text = element_text(size = 12), # legend text size
        legend.title = element_text(size = 12), # label legend text size
        axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))

@@

```{r save beta diversity plots}
# Save
tiff("../../Outputs/plots/microbiome/BetaDiversity_PCoAWU_TMAOmdn_2112.tiff", width = 6, height = 4, units = 'in', res = 300)
betaplot_WU
dev.off()
```

@@

```{r Save tertile prevalence csv}
write.csv(prevalence_tert1, "../../Outputs/tables/Microbiome/Prevalence_Genus_Tertile1.csv")
write.csv(prevalence_tert2, "../../Outputs/tables/Microbiome/Prevalence_Genus_Tertile2.csv")
write.csv(prevalence_tert3, "../../Outputs/tables/Microbiome/Prevalence_Genus_Tertile3.csv")
```

@@

```{r Get the microbiome data ready to combine with pheno data}
da <- read.csv("../../Outputs/tables/Microbiome/DESeq2_PSOtmao1G_TMAOTertile.csv", row.names = 1)
resOTU <- rownames(da)
otuData <- as.data.frame(otu_table(PSOtmao1G))
otuData_desRes <- otuData[rownames(otuData) %in% resOTU,]
otuData_desRes <- as.data.frame(t(otuData_desRes))
# rename for ease downstream
names(otuData_desRes)[names(otuData_desRes) == "bb1b75f41ff9c9db1d1de41e8388eb52"] <- "f__Bacteroidaceae_g__Bacteroides"
names(otuData_desRes)[names(otuData_desRes) == "b180a2455b236c103cf0bfe620095736"] <- "f__Peptostreptococcaceae_g__"
names(otuData_desRes)[names(otuData_desRes) == "e1a2800b24cdf9779b28dc897cddb12a"] <- "f__Christensenellaceae_g__"
names(otuData_desRes)[names(otuData_desRes) == "8f5ceded1cd7e86b8271d3fab24d322a"] <- "f__Lachnospiraceae_g__Butyrivibrio"
```

@@

f__Bacteroidaceae_g__Bacteroides + f__Peptostreptococcaceae_g__ + f__Christensenellaceae_g__ + f__Lachnospiraceae_g__Butyrivibrio

```{r Cardiometabolic + Microbiome Multiple Linear Regression}
# Transfored variable list
transformed_keep_cardio_tbl <- c("sex", "age","bmi_final", "ethnicity", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al75", "hdl_bd1", "ldl_bd1", "chol_bd1", "tg_bd1_ln", "glc_bd1_ln","tnfa_bd1_ln", "il6_bd1_ln", "crp_bd1_ln", "cystatinc_bd1", "choline", "betaine", "carnitine","tmao_log", "f__Bacteroidaceae_g__Bacteroides", "f__Peptostreptococcaceae_g__", "f__Christensenellaceae_g__", "f__Lachnospiraceae_g__Butyrivibrio")

#~~~~~~~~
# TMAO  
#~~~~~~~~
# Make a smaller df from master (the old data)
phen_sub <- phen[, colnames(phen) %in% transformed_keep_cardio_tbl]
cvd_biomark <- subset(phen_sub, select = -c(sex, age, ethnicity, cystatinc_bd1, tmao_log, f__Bacteroidaceae_g__Bacteroides, f__Peptostreptococcaceae_g__, f__Christensenellaceae_g__, f__Lachnospiraceae_g__Butyrivibrio)) # change variable here
n <- ncol(cvd_biomark)

@@

#~~~~~~~~
# TMAO + Bug
#~~~~~~~~
# Make a smaller df from master (the old data)
phen_sub <- phen[, colnames(phen) %in% transformed_keep_cardio_tbl]
cvd_biomark <- subset(phen_sub, select = -c(sex, age, ethnicity, cystatinc_bd1, tmao_log, f__Bacteroidaceae_g__Bacteroides, f__Peptostreptococcaceae_g__, f__Christensenellaceae_g__, f__Lachnospiraceae_g__Butyrivibrio)) # change variable here
n <- ncol(cvd_biomark)

# LCMS method
# Run linear model
my_lm <- lapply(1:n, function(x) lm(cvd_biomark[,x] ~ tmao_log + sex + age + cystatinc_bd1 + f__Bacteroidaceae_g__Bacteroides + f__Peptostreptococcaceae_g__ + f__Christensenellaceae_g__ + f__Lachnospiraceae_g__Butyrivibrio, phen_sub )) # add bugs

@@

@@

```{r Christensenellaceae genus and BMI}
summary(lm(f__Christensenellaceae_g__ ~ bmi_final + sex + age, phen))
```

>>>>

**Updates:**
On 01-19-2022 I reran many of the analyses with the upadted phyloseq object PSOtmao_NA to see if my results were consistent. I saved this markdown to the update-taxonomy-table branch of the GitHub FL100_TMAO project. Background - When MK was reading my manuscript, she noticed that I have some entries in the taxonomy table that say NA and some that say "g__" or "f__". While these may have different meanings, she suggests simply changing all cases to NA. (Another solution, you could manually differentaite "g__" to "g__ASV1" so that the glomming isn't wrong).

@@

```{r Load phyloseq object}
#load("../../data/processed/microbiome/phyloseq_objects/PSOtmao_211108.RData") #PSOtmao
load(file = "../../data/processed/microbiome/phyloseq_objects/PSOtmao_220119.RData") #PSOtmao_NA
PSOtmao <- PSOtmao_NA
```

@@

# Save
tiff("../../Outputs/plots/microbiome/RelAbund_Fam_TMAOTertiles_2201.tiff", width = 5, height = 4, units = 'in', res = 300)
plot.composition.relAbun
dev.off()
```

@@

```{r DESeq2 analysis - TMAO median}
#~~~~~~~~~~~~~~~
# DESeq analysis
# TMAO median
#~~~~~~~~~~~~~~~
mb <- phyloseq_to_deseq2(PSOtmao1G, ~ sex + age + bmi_final.x + tmao_mdn)
mb.ratio <- DESeq(mb, test = "LRT", reduced = ~ sex + age + bmi_final.x) # remove variable of interest here, TMAO
res.ratio <- results(mb.ratio)
alpha = 0.05
alpha.big = 0.5
sigtab = res.ratio[which(res.ratio$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(PSOtmao1G)[rownames(sigtab), ], "matrix"))
sigtab$family_genus <- paste0(sigtab$Family, sigtab$Genus)
head(sigtab)

@@

results when we look at TMAO median and control for age, sex, and BMI.

@@

# Save
tiff("../../Outputs/plots/microbiome/DESeq2_Genus_TMAOTertile1v3_2201.tiff", width = 8, height = 3, units = 'in', res = 300)
DESeqPlot
dev.off()
```
Save

```{r save deseq2 tmao tertile results}
#write.csv(sigtabgen, "../../Outputs/tables/Microbiome/DESeq2_PSOtmao1G_TMAOTertile_2112.csv")
write.csv(sigtabgen, "../../Outputs/tables/Microbiome/DESeq2_PSOtmao1G_TMAOTertile_220119.csv")
```

@@

```{r save beta matrices}
# Outdated
#save(PCoA_WU, file = "../../data/processed/microbiome/diversity/PSOtmao1G_PCoA_WU.RData")
#save(PCoA_UnWU, file = "../../data/processed/microbiome/diversity/PSOtmao1G_PCoA_UnWU.RData")
#save(PCoA_BC, file = "../../data/processed/microbiome/diversity/PSOtmao1G_PCoA_BC.RData")

#save(dismax_wu, file = "../../data/processed/microbiome/diversity/PSOtmao1G_dismax_WU.RData")
#save(dismax_uwu, file = "../../data/processed/microbiome/diversity/PSOtmao1G_dismax_UWU.RData")
#save(dismax_bc, file = "../../data/processed/microbiome/diversity/PSOtmao1G_dismax_BC.RData")

# With updated taxonomy table and from PSOtmaoNA
save(PCoA_WU, file = "../../data/processed/microbiome/diversity/PSOtmao1G_PCoA_WU_220119.RData")
save(PCoA_UnWU, file = "../../data/processed/microbiome/diversity/PSOtmao1G_PCoA_UnWU_220119.RData")
save(PCoA_BC, file = "../../data/processed/microbiome/diversity/PSOtmao1G_PCoA_BC_220119.RData")

save(dismax_wu, file = "../../data/processed/microbiome/diversity/PSOtmao1G_dismax_WU_220119.RData")
save(dismax_uwu, file = "../../data/processed/microbiome/diversity/PSOtmao1G_dismax_UWU_220119.RData")
save(dismax_bc, file = "../../data/processed/microbiome/diversity/PSOtmao1G_dismax_BC_220119.RData")
```

@@

```{r Load bdiversity calculations}
load(file = "../../data/processed/microbiome/diversity/PSOtmao1G_PCoA_WU_220119.RData")
load(file = "../../data/processed/microbiome/diversity/PSOtmao1G_PCoA_UnWU_220119.RData")
load(file = "../../data/processed/microbiome/diversity/PSOtmao1G_PCoA_BC_220119.RData")

load(file = "../../data/processed/microbiome/diversity/PSOtmao1G_dismax_WU_220119.RData")
load(file = "../../data/processed/microbiome/diversity/PSOtmao1G_dismax_UWU_220119.RData")
load(file = "../../data/processed/microbiome/diversity/PSOtmao1G_dismax_BC_220119.RData")

@@

# ADONIS (perMANOVA) analysis
adonis2(dismax_wu ~ tmao_log + sex + age + bmi_final.y, data=df, method = "wunifrac", perm=999) # .076
#adonis2(tmao_log ~ dismax_wu + sex*age + cystatinc_bd1, data=df, method = "wunifrac", perm=999)

@@

#~~~~~~~~~~~
# BrayCurtis
#~~~~~~~~~~~
# Beta dispersion
bdis <- betadisper(dismax_bc, df$tmao_mdn)
permutest(bdis) # .04
boxplot(bdis, las = 2)

@@

# Unweighted
betaplot_UWU <- plot_ordination(PSOtmao1G, PCoA_UnWU, 
                    color="tmao_mdn",
                    shape = NULL,
                    axes = 1:2) +
  scale_color_brewer(palette = "Dark2", labels = c("Below", "Above")) +
  stat_ellipse(linetype = 1) +
  labs(color = "TMAO Median Threshold") +
  theme_minimal() + 
  theme(axis.text = element_text(size = 12), # x and y axis text size
        axis.title = element_text(size = 12), # x and y label text size
        legend.text = element_text(size = 12), # legend text size
        legend.title = element_text(size = 12), # label legend text size
        axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))

@@

```{r save beta diversity plots}
# Save
#tiff("../../Outputs/plots/microbiome/BetaDiversity_PCoAWU_TMAOmdn_2112.tiff", width = 6, height = 4, units = 'in', res = 300)
tiff("../../Outputs/plots/microbiome/BetaDiversity_PCoA_WU_TMAOmdn_2201.tiff", width = 6, height = 4, units = 'in', res = 300)
betaplot_WU
dev.off()

# unweighted
#tiff("../../Outputs/plots/microbiome/BetaDiversity_PCoAWU_TMAOmdn_2112.tiff", width = 6, height = 4, units = 'in', res = 300)
tiff("../../Outputs/plots/microbiome/BetaDiversity_PCoA_UWU_TMAOmdn_2201.tiff", width = 6, height = 4, units = 'in', res = 300)
betaplot_UWU
dev.off()
```

@@

```{r Save Genus tertile prevalence csv}
write.csv(prevalence_tert1, "../../Outputs/tables/Microbiome/Prevalence_Genus_Tertile1_2201.csv")
write.csv(prevalence_tert2, "../../Outputs/tables/Microbiome/Prevalence_Genus_Tertile2_2201.csv")
write.csv(prevalence_tert3, "../../Outputs/tables/Microbiome/Prevalence_Genus_Tertile3_2201.csv")
```

# Family Level Analysis

Repeat TMAO analyses at the Family Level

```{r Family Level}
# Agglomerate taxa
PSOtmao1F = tax_glom(PSOtmao1, "Family", NArm = TRUE)

# Now subset at the genus level to identify the most common genus
# Subset by tertiles (Genus)
PSOtmao1F_tertile1 <- subset_samples(PSOtmao1F, tmao_tertile == "Tertile1")
PSOtmao1F_tertile2 <- subset_samples(PSOtmao1F, tmao_tertile == "Tertile2")
PSOtmao1F_tertile3 <- subset_samples(PSOtmao1F, tmao_tertile == "Tertile3")

#~~~~~~~~~~
# Tertile 1
#~~~~~~~~~~
# Prevalence - how many samples the taxa is observed in
prevalenceF_tert1 = apply(X = otu_table(PSOtmao1F_tertile1),
               MARGIN = ifelse(taxa_are_rows(PSOtmao1F_tertile1), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})

prevalenceF_tert1 = data.frame(Prevalence = prevalenceF_tert1,
                    TotalAbundance = taxa_sums(PSOtmao1F_tertile1), # the total number of occurrences (even if only observed 50 times in 1 sample, the number would be 50)
                    tax_table(PSOtmao1F_tertile1))

#~~~~~~~~~~
# Tertile 2
#~~~~~~~~~~
# Prevalence - how many samples the taxa is observed in
prevalenceF_tert2 = apply(X = otu_table(PSOtmao1F_tertile2),
               MARGIN = ifelse(taxa_are_rows(PSOtmao1F_tertile2), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})

prevalenceF_tert2 = data.frame(Prevalence = prevalenceF_tert2,
                    TotalAbundance = taxa_sums(PSOtmao1F_tertile2), # the total number of occurrences (even if only observed 50 times in 1 sample, the number would be 50)
                    tax_table(PSOtmao1F_tertile2))
#~~~~~~~~~~
# Tertile 3
#~~~~~~~~~~
# Prevalence - how many samples the taxa is observed in
prevalenceF_tert3 = apply(X = otu_table(PSOtmao1F_tertile3),
               MARGIN = ifelse(taxa_are_rows(PSOtmao1F_tertile3), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})

prevalenceF_tert3 = data.frame(Prevalence = prevalenceF_tert3,
                    TotalAbundance = taxa_sums(PSOtmao1F_tertile3), # the total number of occurrences (even if only observed 50 times in 1 sample, the number would be 50)
                    tax_table(PSOtmao1F_tertile3))
```

```{r Save Family tertile prevalence csv}
write.csv(prevalenceF_tert1, "../../Outputs/tables/Microbiome/Prevalence_Family_Tertile1_2201.csv")
write.csv(prevalenceF_tert2, "../../Outputs/tables/Microbiome/Prevalence_Family_Tertile2_2201.csv")
write.csv(prevalenceF_tert3, "../../Outputs/tables/Microbiome/Prevalence_Family_Tertile3_2201.csv")
```

```{r DESeq2 analysis Family - TMAO tertile}
# DESeq analysis - tmao_tertile
mb <- phyloseq_to_deseq2(PSOtmao1F, ~ sex + age + bmi_final.x + tmao_tertile)
mb.ratio <- DESeq(mb, test = "LRT", reduced = ~ sex + age + bmi_final.x) # remove variable of interest here, TMAO
res.ratio <- results(mb.ratio, contrast = c("tmao_tertile", "Tertile1", "Tertile3")) # Look at most different tertiles
alpha = 0.05
sigtab = res.ratio[which(res.ratio$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(PSOtmao1F)[rownames(sigtab), ], "matrix"))
sigtab$order_family <- paste0(sigtab$Order, sigtab$Family)
head(sigtab)

#~~~~~~~~~~~~~~~
# Plot Results
#~~~~~~~~~~~~~~~
resOTU <- rownames(sigtab)
otuData <- as.data.frame(otu_table(PSOtmao1F))
otuData_desRes <- otuData[rownames(otuData) %in% resOTU,]
#otuData_desRes <- as.data.frame(t(otuData_desRes))
# Get sample data groupings
sampleData <- as.data.frame(sample_data(PSOtmao1F))
sampleData <- subset(sampleData, select = c(tmao_mdn, tmao_quantile, tmao_tertile, tmao))
# Get tax table
taxaData <- as.data.frame(tax_table(PSOtmao1F))
taxaData_res <- taxaData[rownames(taxaData) %in% resOTU,]
taxaData_res$OrderFamily <- paste0(taxaData_res$Order, taxaData_res$Family)

# Merge
ggData <- merge(taxaData_res, otuData_desRes, by = 0)
rownames(ggData) <-  ggData$OrderFamily
ggData <- ggData[,-(1:9)]
ggData <- as.data.frame(t(ggData))
ggData <- merge(ggData, sampleData, by = 0)
rownames(ggData) <- ggData[,"Row.names"]
ggData <- subset(ggData, select = -c(Row.names))

# Loop Plot
nres <- length(resOTU)
for (i in 1:nres) { 
    print(ggplot(ggData, aes(x = tmao_tertile, y = log(ggData[,i]))) +
            geom_boxplot(aes(color=tmao_tertile)) +
            geom_point() +
            theme_minimal() +
            ylab(paste0("Log of", colnames(ggData[i]))) +
            xlab("TMAO Tertile") +
            scale_color_brewer(palette = "Dark2") +
            labs(color = "TMAO Tertile")
    )
}

# Loop Plot
nres <- length(resOTU)
for (i in 1:nres) { 
    temp_plot <- ggplot(ggData, aes(x = tmao_tertile, y = log(ggData[,i]))) +
            geom_boxplot(aes(color=tmao_tertile)) +
            geom_point() +
            theme_minimal() +
            ylab(paste0("Log of", colnames(ggData[i]))) +
            xlab("TMAO Tertile") +
            scale_color_brewer(palette = "Dark2") +
            labs(color = "TMAO Tertile")
    tiff(paste0("../../Outputs/plots/microbiome/DESeq2_Family_TMAOTertile_Boxplot_", i,"_2201.tiff"), width = 4, height = 4, units = 'in', res = 300)
    print(temp_plot)
    dev.off()
}

#~~~~~~~~~~~~~~~
# Plot Results Fold Change
#~~~~~~~~~~~~~~~
sigtabgen = subset(sigtab, !is.na(Family))
# Phylum order
x = tapply(sigtabgen$log2FoldChange, sigtabgen$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtabgen$Phylum = factor(as.character(sigtabgen$Phylum), levels=names(x))
# Genus order
x = tapply(sigtabgen$log2FoldChange, sigtabgen$Family, function(x) max(x))
x = sort(x, TRUE)
sigtabgen$Family = factor(as.character(sigtabgen$Family), levels=names(x))
DESeqPlot <- ggplot(sigtabgen, aes(y=order_family, x=log2FoldChange, color=Phylum)) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3) + 
  ylab("Taxonomy") +
  xlab("Log2 Fold Change") + 
  scale_color_brewer(palette = "Paired") +
  theme_minimal() +
  theme(axis.text = element_text(size = 12), # x and y axis text size
        axis.title = element_text(size = 12), # x and y label text size
        legend.text = element_text(size = 12), # legend text size
        legend.title = element_text(size = 12), # label legend text size
        axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))

# Save
tiff("../../Outputs/plots/microbiome/DESeq2_Family_TMAOTertile1v3_2201.tiff", width = 8, height = 3, units = 'in', res = 300)
DESeqPlot
dev.off()
```
Save

```{r save deseq2 Family tmao tertile results}
#write.csv(sigtabgen, "../../Outputs/tables/Microbiome/DESeq2_PSOtmao1G_TMAOTertile_2112.csv")
write.csv(sigtabgen, "../../Outputs/tables/Microbiome/DESeq2_PSOtmao1F_TMAOTertile_220120.csv")
```

@@

```{r Get the genus microbiome data ready to combine with pheno data}
da <- read.csv("../../Outputs/tables/Microbiome/DESeq2_PSOtmao1G_TMAOTertile_220119.csv", row.names = 1)
taxaName_famgen <- da[,"family_genus", drop = F]
taxaName_famgen
resOTU <- rownames(da)
otuData <- as.data.frame(otu_table(PSOtmao1G))
otuData_desRes <- otuData[rownames(otuData) %in% resOTU,]
otuData_desRes <- as.data.frame(t(otuData_desRes))
# rename for ease downstream
#names(otuData_desRes)[names(otuData_desRes) == "bb1b75f41ff9c9db1d1de41e8388eb52"] <- "f__Bacteroidaceae_g__Bacteroides"
#names(otuData_desRes)[names(otuData_desRes) == "b180a2455b236c103cf0bfe620095736"] <- "f__Peptostreptococcaceae_g__"
#names(otuData_desRes)[names(otuData_desRes) == "e1a2800b24cdf9779b28dc897cddb12a"] <- "f__Christensenellaceae_g__"
#names(otuData_desRes)[names(otuData_desRes) == "8f5ceded1cd7e86b8271d3fab24d322a"] <- "f__Lachnospiraceae_g__Butyrivibrio"

# Update PSOtmao_NA results
names(otuData_desRes)[names(otuData_desRes) == "a0798db0c868e9ad6089502956d68d87"] <- "f__Erysipelotrichaceae_g__Catenibacterium"
names(otuData_desRes)[names(otuData_desRes) == "c5859b7f5d14b8c145fbc3ad583c70e8"] <- "f__Erysipelotrichaceae_g__Coprobacillus"
names(otuData_desRes)[names(otuData_desRes) == "5c4ca852b40641b3eb0ad23e69bb6583"] <- "f__Lachnospiraceae_g__Roseburia"
names(otuData_desRes)[names(otuData_desRes) == "8f5ceded1cd7e86b8271d3fab24d322a"] <- "f__Lachnospiraceae_g__Butyrivibrio"
```

@@

Old results:
f__Bacteroidaceae_g__Bacteroides + f__Peptostreptococcaceae_g__ + f__Christensenellaceae_g__ + f__Lachnospiraceae_g__Butyrivibrio

New results:
f__Erysipelotrichaceae_g__Catenibacterium + f__Erysipelotrichaceae_g__Coprobacillus + f__Lachnospiraceae_g__Roseburia + f__Lachnospiraceae_g__Butyrivibrio

```{r Cardiometabolic + Microbiome Multiple Linear Regression}
# Transfored variable list
transformed_keep_cardio_tbl <- c("sex", "age","bmi_final", "ethnicity", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al75", "hdl_bd1", "ldl_bd1", "chol_bd1", "tg_bd1_ln", "glc_bd1_ln","tnfa_bd1_ln", "il6_bd1_ln", "crp_bd1_ln", "cystatinc_bd1", "choline", "betaine", "carnitine","tmao_log", "f__Erysipelotrichaceae_g__Catenibacterium", "f__Erysipelotrichaceae_g__Coprobacillus", "f__Lachnospiraceae_g__Roseburia", "f__Lachnospiraceae_g__Butyrivibrio")

#~~~~~~~~
# TMAO  
#~~~~~~~~
# Make a smaller df from master (the old data)
phen_sub <- phen[, colnames(phen) %in% transformed_keep_cardio_tbl]
#cvd_biomark <- subset(phen_sub, select = -c(sex, age, ethnicity, cystatinc_bd1, tmao_log, f__Bacteroidaceae_g__Bacteroides, f__Peptostreptococcaceae_g__, f__Christensenellaceae_g__, f__Lachnospiraceae_g__Butyrivibrio)) # change variable here
cvd_biomark <- subset(phen_sub, select = -c(sex, age, ethnicity, cystatinc_bd1, tmao_log, f__Erysipelotrichaceae_g__Catenibacterium, f__Erysipelotrichaceae_g__Coprobacillus, f__Lachnospiraceae_g__Roseburia, f__Lachnospiraceae_g__Butyrivibrio)) # change variable here
n <- ncol(cvd_biomark)

@@

#~~~~~~~~
# TMAO + Bug
#~~~~~~~~
# Make a smaller df from master (the old data)
phen_sub <- phen[, colnames(phen) %in% transformed_keep_cardio_tbl]
#cvd_biomark <- subset(phen_sub, select = -c(sex, age, ethnicity, cystatinc_bd1, tmao_log, f__Bacteroidaceae_g__Bacteroides, f__Peptostreptococcaceae_g__, f__Christensenellaceae_g__, f__Lachnospiraceae_g__Butyrivibrio)) # change variable here
cvd_biomark <- subset(phen_sub, select = -c(sex, age, ethnicity, cystatinc_bd1, tmao_log, f__Erysipelotrichaceae_g__Catenibacterium, f__Erysipelotrichaceae_g__Coprobacillus, f__Lachnospiraceae_g__Roseburia, f__Lachnospiraceae_g__Butyrivibrio)) # change variable here
n <- ncol(cvd_biomark)

# LCMS method
# Run linear model
my_lm <- lapply(1:n, function(x) lm(cvd_biomark[,x] ~ tmao_log + sex + age + cystatinc_bd1 + f__Erysipelotrichaceae_g__Catenibacterium + f__Erysipelotrichaceae_g__Coprobacillus + f__Lachnospiraceae_g__Roseburia + f__Lachnospiraceae_g__Butyrivibrio, phen_sub )) # add bugs

@@

Save for supplemental material
```{r Save linear regression results with bacteria}
write.csv(res_des, "../../Outputs/tables/Microbiome/LinReg_cardiometabolic_withDESeq2results_2201.csv")
saveRDS(res_des, "../../Outputs/tables/Microbiome/LinReg_cardiometabolic_withDESeq2results_2201.Rds")
```

Is TNF-a related to the gut microbiome directly?
## Factored by tnfa Tertile

THIS CHUNK WORK IN PROGRESS

```{r DESeq2 analysis - tnf-a tertile}
# Quick check
summary(lm(tnfa_bd1_ln ~ f__Erysipelotrichaceae_g__Catenibacterium, phen_sub)) # NS
summary(lm(tnfa_bd1_ln ~ f__Erysipelotrichaceae_g__Coprobacillus, phen_sub)) # NS
summary(lm(tnfa_bd1_ln ~ f__Lachnospiraceae_g__Roseburia, phen_sub)) # NS
summary(lm(tnfa_bd1_ln ~ f__Lachnospiraceae_g__Butyrivibrio, phen_sub)) # NS
# With covariates
summary(lm(tnfa_bd1_ln ~ f__Erysipelotrichaceae_g__Catenibacterium + sex + age + bmi_final, phen_sub)) # NS
summary(lm(tnfa_bd1_ln ~ f__Erysipelotrichaceae_g__Coprobacillus + sex + age + bmi_final, phen_sub)) # NS
summary(lm(tnfa_bd1_ln ~ f__Lachnospiraceae_g__Roseburia + sex + age + bmi_final, phen_sub)) # NS
summary(lm(tnfa_bd1_ln ~ f__Lachnospiraceae_g__Butyrivibrio + sex + age + bmi_final, phen_sub)) # NS

@@

```{r Christensenellaceae family and BMI}
# Note = ggData is the family level DESeq results 
fam_res <- merge(ggData, phen, by = 0)
rownames(fam_res) <- fam_res[,"Row.names"]
fam_res <- fam_res[,-1]
colnames(fam_res) <- gsub(" ", "_", colnames(fam_res))

summary(lm(`_o__Clostridiales_f__Christensenellaceae` ~ bmi_final + sex + age, fam_res))
```
